Clustering and Parameter Estimation for the Popularity Adjusted Block Model

### 2020-07-22 summary.Rmd
still in the process of moving simulations to separate scripts

```{r}
import::from(magrittr, `%>%`)
library(plot.matrix)
library(ggplot2)
```

@@

```{r}
set.seed(123456)

@@

```{r xu, fig.cap = 'Latent positions $XU$ used to construct the PABM edge probability matrix $P$ under the GRDPG framework. Note that the the points follow the properties outlined in Section 3.2.'}
perm <- cbind(c(1, 0, 0, 0),
              c(0, 0, 1, 0),
              c(0, 1, 0, 0),
              c(0, 0, 0, 1))
U <- eigen(perm)$vectors[, c(3, 1, 2, 4)]
XU <- X %*% U

@@

```{r ase_of_p, fig.cap = 'The adjacency spectral embedding of the edge probability matrix $P$. Note that even though $P = (XU) I_{p, q} (XU)^\\top$, the resulting embedding does not follow the same properties of $XU$, even up to rotation.'}
pairs(Z, asp = 1, col = z)
```

@@

```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.', cache = TRUE}
XU.prime <- transform.embedding(Z, z, runif(4, -pi, pi))
pairs(XU.prime, col = z, asp = 1)
```

@@

```{r a_ase, fig.cap = 'Pair plots of the adjacency spectral embedding of $A$ Since $K=2$ is fixed, we use the signature $(3, 1)$ in this embedding.'}
pairs(Zhat, col = z, asp = 1)
```

@@

```{r q_from_a, cache = TRUE, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).'}
XU.hat <- transform.embedding(Zhat, z, runif(4, -pi, pi))
pairs(XU.hat, col = z, asp = 1)

@@

```{r}
estimate.lambda.block <- function(P.block, within = FALSE) {
  if (within) {
    P.svd <- svd(P.block)
    u <- P.svd$u[, 1]
    if (mean(u) < 0) u <- -u
    s <- sqrt(P.svd$d[1])
    return(s * u)
  } else {
    P.svd <- svd(P.block)
    u <- P.svd$u[, 1]
    if (mean(u) < 0) u <- -u
    v <- P.svd$v[, 1]
    if (mean(v) < 0) v <- -v
    s <- sqrt(P.svd$d[1])
    return(list(s * u, s * v))
  }
}

lambda.rmse <- function(lambda.matrix, Phat, clustering) {
  K <- max(clustering)
  n <- nrow(Phat)
  n.vector <- sapply(seq(K), function(k) sum(clustering == k))
  sapply(seq(K), function(k) {
    low.ind.k <- ifelse(k == 1, 1, sum(n.vector[seq(k - 1)]) + 1)
    high.ind.k <- sum(n.vector[seq(k)])
    sapply(seq(k), function(l) {
      P.block <- Phat[clustering == k, clustering == l]
      if (k == l) {
        lambda <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
        lambda.hat <- estimate.lambda.block(P.block, TRUE)
        return(sum((lambda - lambda.hat) ** 2))
      } else {
        low.ind.l <- ifelse(l == 1, 1, sum(n.vector[seq(l = 1)]) + 1)
        high.ind.l <- sum(n.vector[seq(l)])
        lambda.kl <- as.numeric(lambda.matrix[seq(low.ind.k, high.ind.k), l])
        lambda.lk <- as.numeric(lambda.matrix[seq(low.ind.l, high.ind.l), k])
        lambda.hats <- estimate.lambda.block(P.block)
        return(sum((lambda.kl - lambda.hats[[1]]) ** 2) + 
                 sum((lambda.lk - lambda.hats[[2]]) ** 2))
      }
    }) %>% 
      sum() %>% 
      return()
  }) %>% 
    sum() %>% 
    magrittr::divide_by(n * K) %>% 
    sqrt() %>% 
    return()
}
```

@@

```{r lambda_est, cache = TRUE, fig.cap = 'Median and IQR RMSE from Algorithm 4 using sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.'}
rmse.df <- readr::read_csv('rmse.csv')

>>>>

```{r}
import::from(magrittr, `%>%`)
library(plot.matrix)
library(ggplot2)

theme_set(theme_bw())
```

@@

```{r eval = FALSE}
set.seed(123456)

@@

```{r xu, fig.cap = 'Latent positions $XU$ used to construct the PABM edge probability matrix $P$ under the GRDPG framework. Note that the the points follow the properties outlined in Section 3.2.', eval = FALSE}
perm <- cbind(c(1, 0, 0, 0),
              c(0, 0, 1, 0),
              c(0, 1, 0, 0),
              c(0, 0, 0, 1))
U <- eigen(perm)$vectors[, c(3, 1, 2, 4)]
XU <- X %*% U

@@

```{r ase_of_p, fig.cap = 'The adjacency spectral embedding of the edge probability matrix $P$. Note that even though $P = (XU) I_{p, q} (XU)^\\top$, the resulting embedding does not follow the same properties of $XU$, even up to rotation.', eval = FALSE}
pairs(Z, asp = 1, col = z)
```

@@

```{r q_trans, fig.cap = 'Pair plots of $ZQ$ where $Q$ was found by minimizing objective function (1). This embedding is equivalent to $XU$ up to 90-degree rotations and reflection along axes.', cache = TRUE, eval = FALSE}
XU.prime <- transform.embedding(Z, z, runif(4, -pi, pi))
pairs(XU.prime, col = z, asp = 1)
```

@@

```{r a_ase, fig.cap = 'Pair plots of the adjacency spectral embedding of $A$ Since $K=2$ is fixed, we use the signature $(3, 1)$ in this embedding.', eval = FALSE}
pairs(Zhat, col = z, asp = 1)
```

@@

```{r q_from_a, cache = TRUE, fig.cap = 'Pair plots of $\\hat{Z} $ where $\\hat{Z}$ is the ASE of an adjacency matrix of a PABM and $Q$ is estimated by minimizing objective function (1).', eval = FALSE}
XU.hat <- transform.embedding(Zhat, z, runif(4, -pi, pi))
pairs(XU.hat, col = z, asp = 1)

@@

@@

```{r lambda_est, cache = FALSE, fig.cap = 'Median and IQR RMSE from Algorithm 4 using sample sizes from 64 to 1024. Simulations were repeated 100 times for each sample size.'}
rmse.df <- readr::read_csv('rmse.csv')
